Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''

### 2024-08-27 etm-transform/manuscript/main.tex
Added experimental data

When instantiating an instance $\kem(\keygen, \encap, \decap)$, there are a variety of possible MAC's to choose from. For each of the security level and choice of MAC, we measured the number of CPU cycles needed to produce a digest of Kyber's ciphertext. The MAC implementations are taken from OpenSSL and run on an Intel MacBook Pro with i7-9700h (TODO: run it on a server). The median measurements are reported in table \ref{tbl:standalone-mac-perf}.

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        Name & Security & 768 bytes & 1088 bytes & 1568 bytes \\
        \hline
        CMAC & many-time & 5022 & 5442 & 6090 \\
        \hline
        KMAC-256 & many-time & 7934 & 9862 & 11742 \\
        \hline
        GMAC & one-time & 2778 & 2756 & 2762 \\
        \hline
        Poly1305 & one-time & 1128 & 1218 & 1338 \\
        \hline
    \end{tabular}
    \caption{Standalone $\mac$ performances}\label{tbl:standalone-mac-perf}
\end{table}

Based on the security reduction in section \ref{sec:main-results} we chose standalone GMAC and Poly1305 for their substantial performance advantage over other constructions. We modified the reference implementation (https://github.com/pq-crystals/kyber) to implement the proposed construction, then measured the CPU cycles needed to run each of routines. The median CPU cycles are reported in table \ref{tbl:kyber-ae-perf}.

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        Name & Security level & $\keygen$ & $\encap$ & $\decap$ \\
        \hline
        Kyber512 & 128 bits &  & &  \\
        \hline
        Kyber768 & 192 bits &  & &  \\
        \hline
        Kyber1024 & 256 bits &  & &  \\
        \hline
        Kyber512 + GMAC& 128 bits &  & &  \\
        \hline
        Kyber768 + GMAC& 192 bits &  & &  \\
        \hline
        Kyber1024 + GMAC& 256 bits &  & &  \\
        \hline
        Kyber512 + Poly1305& 128 bits &  & &  \\
        \hline
        Kyber768 + Poly1305& 192 bits &  & &  \\
        \hline
        Kyber1024 + Poly1305& 256 bits &  & &  \\
        \hline
    \end{tabular}
    \caption{Performance measurements}\label{tbl:kyber-ae-perf}
\end{table}

>>>>

When instantiating an instance of $\kem(\keygen, \encap, \decap)$, there are a variety of MAC's to chose from. We experimented with Poly1305, AES-256-GCM (aka GMAC), AES-256-CBC (aka CMAC), and KMAC-256. We instantiated an intance of the ``encrypt-then-MAC'' KEM using each of the chosen MAC, then measured the median number of CPU cycles needed to perform the key generation, encapsulation, and decapsulation routines among 10000 runs. The Kyber implementation is taken from the reference implementation (https://github.com/pq-crystals/kyber). MAC implementations are taken from OpenSSL 3.3.1. The source code is compiled with GCC 11.4.1 on Amazon Linux 2. Performance measurements were taken from a c7a.medium AWS EC2 instance with a AMD EPYC 9R14 (1) @ 3.700GHz. The experimental results are listed in table \ref{tbl:kyber-ae-perf}.

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        Name & Security level & $\keygen$ & $\encap$ & $\decap$ \\
        \hline
        Kyber512 & 128 bits & 75945 & 91467& 121185 \\
        \hline
        Kyber512 + Poly1305& 128 bits & 76907 & 93157 & 33733 \\
        \hline
        Kyber512 + GMAC& 128 bits & 76917 & 95419 & 37725 \\
        \hline
        Kyber512 + CMAC& 128 bits & 76907 & 99839 & 40117 \\
        \hline
        Kyber512 + KMAC-256& 128 bits & 76387 & 101009 & 40741 \\
        \hline
        Kyber768 & 192 bits & 129895 & 146405& 186445 \\
        \hline
        Kyber768 + Poly1305& 192 bits & 128205 & 146405& 43315 \\
        \hline
        Kyber768 + GMAC& 192 bits & 127997 & 149525 & 46513 \\
        \hline
        Kyber768 + CMAC& 192 bits & 129167 & 151007 & 49841 \\
        \hline
        Kyber768 + KMAC-256& 192 bits & 128829 & 155219 & 52415 \\
        \hline
        Kyber1024 & 256 bits & 194921 & 199185 & 246245 \\
        \hline
        Kyber1024 + Poly1305& 256 bits & 196013 & 205763 & 51375 \\
        \hline
        Kyber1024 + GMAC& 256 bits & 196039 & 208805 & 54573 \\
        \hline
        Kyber1024 + CMAC& 256 bits & 195389 & 213667 & 59175 \\
        \hline
        Kyber1024 + KMAC-256& 256 bits & 196117 & 216761 & 62269 \\
        \hline
    \end{tabular}
    \caption{Performance measurements}\label{tbl:kyber-ae-perf}
\end{table}
